---
title: "Surv_analysis"
author: "Uche"
date: "2024-03-21"
output: html_document
---

#### Proton pump inhibitor(PPI) VS Histamine H2-receptor antagonist(H2RA); all cause mortality

#### Install and load the required libraries

```{r}
# install commands have been commented out, install only if packages are missing. Libraries can be loaded if missing too.
#install.packages("tidyverse")
#install.packages("survival")
#install.packages("table1")
#install.packages("survminer")
#install.packages("knitr")
#install.packages("kableExtra")
#install.packages("broom")

#load required libraries if missing
library(tidyverse)
library(survival)
library(table1)
library(survminer)
library(consort)
library(knitr)
library(kableExtra)
library(broom)

```

### 1. Load data and codelists into two complied list objects

```{r}
# Lists files directory matching the specified pattern, subset string to lowercase and remove unneeded characters

files <- list.files("Assessment_sample", full.names = TRUE, pattern = ".csv")
filenames <- str_to_lower(str_remove(str_remove(tools::file_path_sans_ext(basename(files)), "_assessment_24"), "Sim_"))

# Apply the read_csv function to all files and save the output as one big list
data <- lapply(files, readr::read_csv)
names(data) <- filenames
sapply(data, function(x) sapply(x, class))
# Change some varaiables to factor/categorical
# Specify which columns should be factors
factor_vars <- c("gender", "eth5", "constype", "imd_person")
# For all datasets, across any of the above specified columns, convert to factors
data <- map(data, ~.x %>% mutate(across(any_of(factor_vars), as_factor)))


## For the codelist
# Make a list of all file paths and file names
files <- list.files("Codelists", full.names = TRUE, pattern = "codes.csv")
filenames <- str_to_lower(str_remove(tools::file_path_sans_ext(basename(files)), "_codes"))

# Apply the read_csv function to all files and save the output as one big list
codes <- lapply(files, readr::read_csv)
names(codes) <- filenames
```

#### Initial exploratory analysis

```{r}
#Reorder deprivation status
levels(data$imd$imd_person)
data$imd$imd_person <- factor(data$imd$imd_person, levels = c("Least Deprived (1)", "2", "3", "4", "Most Deprived (5)"))
levels(data$imd$imd_person)

#Chech total number of unique patients
data$patient %>%
  summary()
data$patient %>%
  summarise(n_rows = n(),
            n_patients = n_distinct(patid))

```

#### Combine PPI and H2RA prescriptions and create cohort based on 1st prescription

```{r}
#Inspect ppi and h2ra codelists
data$therapy %>% 
  summary()
head(codes$ppi)
head(codes$h2ra)

# Extract all PPI prescriptions, join with therapy dataset on prodcode
ppis <- data$therapy %>% 
  inner_join(codes$ppi, by="prodcode") %>%
  mutate(ppi=1)

# Extract all H2RA prescriptions, join with therapy dataset on prodcode
h2ras<- data$therapy %>% 
  inner_join(codes$h2ra, by="prodcode") %>%
  mutate(ppi=0)

# Combine PPI and H2RA prescriptions into one dataframe
ppis_h2ras<-rbind(ppis, h2ras)

cohort <- ppis_h2ras %>%  # Take the list of PPI/H2RA prescriptions
  arrange(patid, eventdate) %>% # Sort by patient, then by prescription date
  filter(!duplicated(patid)) # Take the top row per patient (i.e. earliest prescription date)
nrow(cohort)
```

#### Create cohort and implement two of the eligibility criteria

-   After the date of the patient's registration at the general practice plus one year

-   Between 17 April 1997 and 17 April 2017

```{r}
# Take the cohort dataset and join the "Patient" dataset
# keep only observations where the eventdate is later than the registration date + 1 year

#Criteria 2
cohort <- cohort %>% 
  left_join(data$patient[c("patid", "crd")], by="patid") %>% 
  filter(eventdate > crd+365.25)
nrow(cohort)

#Criteria 3
# Filter cohort to contain observations 17 April 1997 and 17 April 2017
cohort <- cohort %>% 
  filter(eventdate >= as.Date("1997-04-17") & eventdate <= as.Date("2017-04-17"))
nrow(cohort)

# Create flowchart dataset with one row per patient, with a variable indicating reason for exclusion
flowchart <- ppis_h2ras %>%   
  arrange(patid, eventdate) %>%  
  filter(!duplicated(patid))  %>% 
  left_join(data$patient[c("patid", "crd", "eth5")], by="patid") %>%  
  mutate(excluded = case_when((eventdate < crd+365.25) ~ "Insufficient registration", 
                              (eventdate < as.Date("1997-04-17") | eventdate > as.Date("2017-04-17")) ~ "Prescription out of dates",
                              eth5 == "Unknown" ~ "Unrecorded ethnicity"
                              )) %>%
  mutate(ppi_str = case_when(ppi==1 ~ "PPI", ppi==0 ~ "H2RA")) 

# Create flowchart
flow <- consort_plot(data = flowchart,
                     orders = c(patid =    "Patients with PPI/H2RA prescription",
                                excluded = "Excluded",
                                ppi_str =  "Received eligible prescription",
                                patid =    "Analysed"),
                     side_box = c("excluded"),
                     allocation = "ppi_str")

# Display flowchart
plot(flow)

#change index date name
cohort <- cohort %>% 
  select(patid, eventdate, ppi) %>% 
  rename(indexdate=eventdate)
cohort
```

#### Extract data for patients with prior history of gastric cancer

```{r}
#Gastric cancer data
prior_gastric_cancer <- data$clinical %>% 
  inner_join(codes$gastric_cancer, by="medcode") %>% 
  arrange(patid, eventdate) %>% 
  filter(!duplicated(patid)) %>% 
  inner_join(cohort, by="patid") %>% 
  filter(eventdate < indexdate) %>% 
  mutate(prior_gastric_cancer=1) %>% 
  select(patid, prior_gastric_cancer)
prior_gastric_cancer
```

#### Extract data for patients with GERD last 6 months prior to prescription

```{r}
#Gerd data
recent_gerd <- data$clinical %>% 
  inner_join(cohort, by="patid") %>% 
  arrange(patid, eventdate) %>% 
  inner_join(codes$gerd, by="medcode") %>% 
  filter(eventdate>=indexdate-180 & eventdate<=indexdate) %>% 
  filter(!duplicated(patid)) %>% 
  select(patid) %>% 
  mutate(recent_gerd=1) 
recent_gerd
```

#### Extract data for patients with peptic ulcer 6 months prior to prescription

```{r}
#Peptic ulcer data
recent_pepticulcer <- data$clinical %>% 
  inner_join(cohort, by="patid") %>% 
  arrange(patid, eventdate) %>% 
  inner_join(codes$peptic_ulcer, by="medcode") %>% 
  filter(eventdate>=indexdate-180 & eventdate<=indexdate) %>% 
  filter(!duplicated(patid)) %>% 
  select(patid) %>% 
  mutate(recent_pepticulcer=1) 
recent_gerd

```

#### Extract data for patients with specific consultations

```{r}
#Consultation data for specific consultations
filtered_consultations <- data$consultations %>%
  filter(constype %in% c("Surgery consultation", "Follow-up/routine visit", "Clinic", "Telephone call from a patient", "Acute visit", "Home Visit", "Emergency Consultation"))%>%
  inner_join(cohort, by="patid") %>% 
  arrange(patid, eventdate) %>% 
  filter(eventdate>=indexdate-365.25 & eventdate<=indexdate) %>% 
  group_by(patid) %>%
  summarize(total_consultations = n())
```

#### Create study end dates for observations

```{r}
# Join cohort with patient data to create enddate by patient id and specific requirements; minimum of the study end date or tod or deathdate
enddates <- cohort %>% 
  left_join(data$patient, by="patid") %>% 
  mutate(enddate=pmin( 
    as.Date("2017-04-17"), 
    pmin(tod, deathdate, na.rm = TRUE),
    na.rm = TRUE),
    died=ifelse(is.na(deathdate), 0, (deathdate==enddate))) %>% 
  select(patid, deathdate, enddate, died)
enddates
```

#### Demographics datasframe

Data frame with cohort and selected specific demographic characteristics, age at first prescription and also a stratified calendar period

```{r}
# Dataset with deprivation status
demographics <- cohort %>% 
  left_join(data$patient, by="patid") %>%
  left_join(data$imd, by="patid") %>% 
  mutate(age=round(as.numeric(indexdate-as.Date(paste0(yob, "-06-15")))/365.25),
         calendarperiod=case_when(indexdate<as.Date("2000-01-01") ~ "1997-1999",
                                  indexdate<as.Date("2005-01-01") ~ "2000-2004",
                                  indexdate<as.Date("2010-01-01") ~ "2005-2009",
                                  indexdate<as.Date("2015-01-01") ~ "2010-2014",
                                  indexdate>=as.Date("2015-01-01") ~ "2015-2017")) %>% 
  rename(pracid=pracid.x) %>%
  select(patid, pracid, age, gender, imd_person,eth5, calendarperiod)
demographics
```

#### Extract BMI data: calculate BMI from raw BMI and also Weight and Height

```{r}
bmi_data <- data$clinical %>% 
  filter(enttype==13) %>% # Keep weight/BMI measurements only
  rename(bmi_date=eventdate) %>%
  inner_join(cohort, by="patid") %>%
  inner_join(data$additional, by= c("patid", "enttype", "adid"))  %>% 
  rename(bmi=data3) %>% # Pick out BMI measurements
  filter(as.numeric(indexdate - bmi_date) >= 0 & as.numeric(indexdate - bmi_date) <= 5 * 365.25) %>% # Keep measurements within 5 years
  filter(!is.na(bmi), between(bmi, 5, 200)) %>% # Delete extreme measurements
  group_by(patid, bmi_date) %>%
  summarize(bmi = mean(bmi)) %>% # Average measurements on same day
  arrange(patid, bmi_date) %>% 
  ungroup() %>%
  group_by(patid)  %>%
  filter(bmi_date == max(bmi_date)) %>% # Keep one measurement (the latest)
  mutate(preference=2) %>% # Mark as less preferred (vs calculated from weight)
  select(patid, bmi, bmi_date, preference) 

# Check that there is one row per patient in the BMI data
length(unique(bmi_data$patid))==nrow(bmi_data)
weight_data <- data$clinical %>% 
  filter(enttype ==13) %>%  # Keep weight/BMI measurements only
  rename(weight_date=eventdate) %>%
  inner_join(cohort, by="patid") %>%
  inner_join(data$additional, by= c("patid", "enttype", "adid"))  %>% 
  rename(weight_kg=data1) %>% # Pick out weight measurements
  filter(as.numeric(indexdate - weight_date) >= 0 & as.numeric(indexdate - weight_date) <= 5 * 365.25)%>% # Keep measurements within 5 years
  filter(!is.na(weight_kg), weight_kg>=20)  %>%  # Delete extreme measurements
  group_by(patid, weight_date)  %>%
  summarize(weight_kg = mean(weight_kg)) %>% # Average measurements on same day
  arrange(patid, weight_date) %>% 
  ungroup() %>%
  group_by(patid)  %>%
  filter(weight_date == max(weight_date)) %>% # Keep one measurement (the latest)
  select(patid, weight_kg, weight_date)

# Check that there is one row per patient in the weight data
length(unique(weight_data$patid))==nrow(weight_data)

# Height measurements
height_data <- data$clinical %>% 
  filter(enttype ==14)  %>% # Keep height measurements only
  rename(height_date=eventdate) %>%
  inner_join(cohort, by="patid") %>%
  inner_join(data$additional, by= c("patid", "enttype", "adid")) %>%
  rename(height_m=data1) %>%
  mutate(yoh = as.numeric(format(height_date, "%Y")))  %>%
  filter(!is.na(height_m), between(height_m, 1.20, 2.15)) %>% # Remove extreme measurements
  group_by(patid, height_date) %>%
  summarize(height_m = mean(height_m))  %>% # Average measurements on same day
  arrange(patid, height_date) %>% 
  ungroup() %>%
  group_by(patid)  %>%
  filter(height_date == max(height_date)) %>% # Keep one measurement (the latest)
  select(patid, height_m, height_date)  

# Check that there is one row per patient in the height data
length(unique(height_data$patid))==nrow(height_data)

# Calculate BMI from weight and height
bmi_calculate <- inner_join(height_data, weight_data, by="patid")  %>%
  mutate(bmi=weight_kg/height_m^2, preference=1) %>% 
  rename(bmi_date=weight_date) %>% # Set date of calculation to date of weight
  select(patid, bmi, bmi_date, preference) 

bmi <- rbind(bmi_calculate, bmi_data)
bmi <- bmi %>%
  arrange(patid, preference) %>% # Put the calculated BMI first (per patient)
  filter(!duplicated(patid)) %>% # Take only the first row per patient 
  select(patid, bmi)

# Check that there is one row per patient in the final BMI data
length(unique(bmi$patid))==nrow(bmi)
```

#### Extract analysis dataset

```{r}
#Join different extracted data by common patient id
analysis_dataset <- cohort %>% 
  left_join(prior_gastric_cancer, by="patid") %>% 
  left_join(recent_gerd, by="patid") %>% 
  left_join(recent_pepticulcer, by="patid") %>% 
  left_join(enddates, by="patid") %>% 
  left_join(demographics, by="patid") %>% 
  left_join(bmi, by="patid") %>% 
  left_join(filtered_consultations, by="patid") %>% 
  mutate(across(.cols = c("prior_gastric_cancer", "recent_gerd", "recent_pepticulcer"),#add 0 to observations with no events
                .fns = ~ ifelse(is.na(.x), 0, 1)))
analysis_dataset

```

#### Create survival/follow up time

Replace "0" observed follow up time with value less that minimum

```{r}
# Create survival time from enddate and indexdate
analysis_dataset <- analysis_dataset %>%
  mutate(survtime = as.numeric(difftime(enddate, indexdate, units = "days")))
# Replace 0 with 0.9 and convert survtime to years
analysis_dataset <- analysis_dataset %>%
  mutate(survtime = ifelse(survtime == 0, 0.9, survtime / 365.25))

```

#### Remove unneeded variables, check levels of categorical variables and further eligibility checks

Remove unknown in ethnicity to meet eligibility criteria

Check for recoreded year of birth and gender

```{r}

levels(analysis_dataset$eth5)
table(analysis_dataset$eth5)


#clean and drop unused levels
analysis_dataset <- analysis_dataset %>%
  filter(eth5 != "Unknown")
analysis_dataset$eth5 <- droplevels(analysis_dataset$eth5)
# drop columns not needed for survival analysis for easier datahandling,(pracid, indexdate, deathdate, enddate)
analysis_dataset <- analysis_dataset %>%
  select(-pracid, -indexdate, -deathdate, -enddate)
#Optionally save dataset
#saveRDS(analysis_dataset, file = "analysis_dataset.Rds") 
```

### 2. Descriptive statistics and preliminary analysis

```{r}
# Range of survival time
range(analysis_dataset$survtime)
# Create plot object for easier manipulation
plotdata <- analysis_dataset
# manipulate variables for tabular presentation, assign labels for the levels
plotdata$died <- 
  factor(plotdata$died, levels=c(0,1),
         labels=c("NO", 
                  "YES"))
plotdata$recent_gerd <- 
  factor(plotdata$recent_gerd , levels=c(0,1),
         labels=c("NO", 
                  "YES"))
plotdata$recent_pepticulcer <- 
  factor(plotdata$recent_pepticulcer , levels=c(0,1),
         labels=c("NO", 
                  "YES"))
plotdata$prior_gastric_cancer <- 
  factor(plotdata$prior_gastric_cancer  , levels=c(0,1),
         labels=c("NO", 
                  "YES"))
plotdata$ppi <- 
 factor(plotdata$ppi  , levels=c(0,1),
         labels=c("H2RA", 
                  "PPI"))
#Change labels for interpretability
label(plotdata$age)       <- "AGE"
label(plotdata$died)        <- "DIED"
label(plotdata$eth5)      <- "ETHNICITY"
label(plotdata$gender) <- "GENDER"
label(plotdata$bmi) <- "BMI"
label(plotdata$prior_gastric_cancer) <- "PRIOR GASTRIC CANCER"
label(plotdata$imd_person) <- "DEPRIVATION STATUS"
label(plotdata$recent_pepticulcer) <- "RECENT PEPTIC ULCER"
label(plotdata$recent_gerd) <- "RECENT GERD"
label(plotdata$total_consultations) <- "TOTAL CONSULTATIONS"
label(plotdata$survtime) <- "SURVIVAL TIME"
label(plotdata$calendarperiod) <- "CALENDAR PERIOD"
#Asiign usnits gor age, BMI and survival time
units(plotdata$age)       <- "years"
units(plotdata$bmi) <- "kg/m2"
units(plotdata$survtime) <- "years"
#set mathematical parameters and rounding arguments
caption  <- "Basic stats"
my.render.cont <- function(x) {
  with(stats.apply.rounding(stats.default(x), digits=2), c("",
                                                           "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}
#plot table
mytable <- table1(~ age+eth5+died+gender+ bmi+ imd_person+prior_gastric_cancer+recent_gerd+recent_pepticulcer+ total_consultations+survtime+calendarperiod|ppi,render.continuous=my.render.cont, data=plotdata,topclass="Rtable1-zebra") 
mytable
```

#### Kaplan-Meier plot to investigate the crude association between exposure status and survival, evidence for a difference between the two survival curves using log rank tests

```{r}
# Plot Kaplan-Meier curves with p-values, risk table, and confidence intervals
km <- survfit(Surv(survtime,died)~factor(ppi),data=analysis_dataset)
#Plot KM curve with risktable and logrank p value
ggsurvplot(km, data = analysis_dataset, pval = TRUE, risk.table = "nrisk_cumevents", conf.int = TRUE, xlab="Time since first prescription(years)", legend.labs=c("H2RA","PPI"), title="KAPLAN MEIER PLOT OF EXPOSURE GROUPS(with pvalue)")

# Perform log-rank test
survdiff(Surv(survtime, died) ~ factor(ppi), data = analysis_dataset)
```

### 3. Analysis using Cox Regression

#### Fit a univariable cox model

```{r}
#Fit univariate cox model
univar_cox<-coxph(Surv(survtime,died)~as.factor(ppi),data=analysis_dataset)
summary(univar_cox)
```

#### Cox multivariable and Martingale residual

```{r}
#Remove missing obseavations(missing BMI)
analysis_dataset <- na.omit(analysis_dataset)
#Cox multivariable model
cox_multivariable <- coxph(Surv(survtime, died) ~ 
                             factor(ppi) + 
                             factor(recent_gerd) + 
                             factor(recent_pepticulcer) + 
                             factor(prior_gastric_cancer) + 
                             factor(gender) + 
                             bmi + 
                             age + 
                             factor(eth5) + 
                             factor(imd_person) + 
                             factor(calendarperiod) + 
                             total_consultations,  
                           data = analysis_dataset)

summary(cox_multivariable)
```

#### Martingale residual tests for continuous variables for multivariable model(age and total consultation time

This is done to consider method of imputation of continuous confounders. BMI already restricted to linear

```{r}
# Initiate residual object
martingale_res <- resid(cox_multivariable, type = "martingale")

#martingale residual plots for age
y_min <- -1 #set limits for y axis for closer residual view
y_max <- 1
par(mfrow=c(1,2)) # initiate 2 by 1 plot area
plot(analysis_dataset$age, martingale_res, xlab = "Age", ylab = "Martingale Residual", 
     main = "Martingale Residuals vs. Age", ylim = c(y_min, y_max))
lowess_line <- lowess(analysis_dataset$age, martingale_res, f = 0.5)  # Set smoother span to 0.5
lines(lowess_line, col = "red", lwd = 2)
abline(h=0,lwd=2,col="grey")


# Martingale plot for total consultations
y_min <- -1
y_max <- 1
plot(analysis_dataset$total_consultations, martingale_res, xlab = "Total Consultations", 
     ylab = "Martingale Residual", main = "Martingale Residuals vs. Total Consultations")
lines(lowess(analysis_dataset$total_consultations, martingale_res),col="red", lwd=2)


```

```         
```

#### Include age interaction term and repeat martingale plot

```{r}
cox_multivariable1 <- coxph(Surv(survtime, died) ~ 
                             factor(ppi) + 
                             factor(recent_gerd) + 
                             factor(recent_pepticulcer) + 
                             factor(prior_gastric_cancer) + 
                             factor(gender) + 
                             bmi + 
                             age + 
                             I(age^3)+
                             factor(eth5) + 
                             factor(imd_person) + 
                             factor(calendarperiod) + 
                             total_consultations,  
                           data = analysis_dataset)

```

#### Refit Martingale residual and replot

```{r}
#Repeat martingale plot
martingale_res2 <- resid(cox_multivariable1, type = "martingale")

y_min <- -1
y_max <- 1
par(mfrow=c(1,1))
plot(analysis_dataset$age, martingale_res2, xlab = "Age", ylab = "Martingale Residual", 
     main = "Martingale Residuals vs. Age", ylim = c(y_min, y_max))
lowess_line <- lowess(analysis_dataset$age, martingale_res2, f = 0.5)  # Set smoother span to 0.5
lines(lowess_line, col = "red", lwd = 2)
abline(h=0,lwd=2,col="grey")



```

#### Assessing proportional hazards assumption using schoenfeld residuals; statistical tests and plots

```{r}
#Initiate residual object on refitted multivariable model
sch.resid<-cox.zph(cox_multivariable1, transform = 'identity')

sch.resid


```

Plot Shoenfeld residuals

```{r}
#plot schoenfeld residuals
plot(sch.resid,col="red",lwd=2)
```

Selected plot of shoenfeld residuals with p value\< 0.05

#### Refit a multiivariable cox model with chosen confounders

```{r}
final_CoxModel <- coxph(Surv(survtime, died) ~ 
                            factor(ppi) + 
                            factor(recent_gerd) +
                            factor(recent_pepticulcer) + 
                            factor(prior_gastric_cancer) + 
                            factor(gender) + 
                            factor(eth5) + 
                            factor(imd_person) +
                            age +
                            I(age^3),
                          data = analysis_dataset)
summary(final_CoxModel)
```

#### Create table object with broom package and knit with kable

```{r}
#Create table to see coefficients of final cox model
coef_table <- broom::tidy( final_CoxModel, exponentiate=TRUE, conf.int=TRUE)
kable(coef_table,caption = 'Coefficients of final cox model', longtable = F) %>%
  kable_styling(font_size = 13) %>% row_spec(0, font_size=12)

```

#### Estimated survivor curves under the two exposure statuses for selected individuals

```{r}
par(mfrow=c(2,2))
# Estimate survival curves for 50yo White male on ppi, no comorbidities
survfit1 <- survfit(final_CoxModel, 
                    newdata = data.frame(ppi = "1", 
                                         age = 50,
                                         gender = "Male",
                                         recent_gerd = "0",
                                         recent_pepticulcer ="0",
                                         prior_gastric_cancer ="0",
                                         eth5 = "White",
                                         imd_person = "Most Deprived (5)"))

# Estimate survival curves for 50yo White male on h2ra, no comorbidities
survfit2 <- survfit(final_CoxModel, 
                    newdata = data.frame(ppi = "0", 
                                         age = 50,
                                         gender = "Male",
                                         recent_gerd = "0",
                                         recent_pepticulcer ="0",
                                         prior_gastric_cancer ="0",
                                         eth5 = "White",
                                         imd_person = "Most Deprived (5)"))

# Plot survival curves
plot(survfit1, col = "blue", lty = 1, xlim = c(0, 20), ylim = c(0, 1), xlab = "Time(years)", ylab = "Survival Probability", main = "50yo White Male,Most Deprived ")
lines(survfit2, col = "red", lty = 1)
legend("bottomleft", legend = c("PPI", "H2RA"), col = c("blue", "red"), lty = 1, cex=0.7)



# Estimate survival curves for 50yo White male on ppi,GERD
survfit3 <- survfit(final_CoxModel, 
                    newdata = data.frame(ppi = "1", 
                                         age = 50,
                                         gender = "Male",
                                         recent_gerd = "1",
                                         recent_pepticulcer ="0",
                                         prior_gastric_cancer ="0",
                                         eth5 = "White",
                                         imd_person = "Least Deprived (1)"))

# Estimate survival curves for 50yo White male on H2RA, GERD
survfit4 <- survfit(final_CoxModel, 
                    newdata = data.frame(ppi = "0", 
                                         age = 50,
                                         gender = "Male",
                                         recent_gerd = "1",
                                         recent_pepticulcer ="0",
                                         prior_gastric_cancer ="0",
                                         eth5 = "White",
                                         imd_person = "Least Deprived (1)"))

# Plot survival curves
plot(survfit3, col = "blue", lty = 1, xlim = c(0, 20), ylim = c(0, 1), xlab = "Time(years)", ylab = "Survival Probability", main = "50yo White Male,GERD,Least Deprived ")
lines(survfit4, col = "red", lty = 1)
legend("bottomleft", legend = c("PPI", "H2RA"), col = c("blue", "red"), lty = 1, cex=0.7)

# Estimate survival curves for 50yo White female on ppi, no comorbidities
survfit5 <- survfit(final_CoxModel, 
                    newdata = data.frame(ppi = "1", 
                                         age = 50,
                                         gender = "Female",
                                         recent_gerd = "0",
                                         recent_pepticulcer ="0",
                                         prior_gastric_cancer ="0",
                                         eth5 = "White",
                                         imd_person = "Most Deprived (5)"))

# Estimate survival curves for 50yo White female on H2RA, no comorbidities
survfit6 <- survfit(final_CoxModel, 
                    newdata = data.frame(ppi = "0", 
                                         age = 50,
                                         gender = "Female",
                                         recent_gerd = "0",
                                         recent_pepticulcer ="0",
                                         prior_gastric_cancer ="0",
                                         eth5 = "White",
                                         imd_person = "Most Deprived (5)"))

# Plot survival curves
plot(survfit5, col = "blue", lty = 1, xlim = c(0, 20), ylim = c(0, 1), xlab = "Time(years)", ylab = "Survival Probability", main = "50yo White Female,Most Deprived ")
lines(survfit6, col = "red", lty = 1)
legend("bottomleft", legend = c("PPI", "H2RA"), col = c("blue", "red"), lty = 1, cex = 0.7)

# Estimate survival curves for 50yo White female on ppi, GERD
survfit7 <- survfit(final_CoxModel, 
                    newdata = data.frame(ppi = "1", 
                                         age = 50,
                                         gender = "Female",
                                         recent_gerd = "1",
                                         recent_pepticulcer ="0",
                                         prior_gastric_cancer ="0",
                                         eth5 = "White",
                                         imd_person = "Least Deprived (1)"))

# Estimate survival curves for 50yo White female on H2RA, GERD
survfit8 <- survfit(final_CoxModel, 
                    newdata = data.frame(ppi = "0", 
                                         age = 50,
                                         gender = "Female",
                                         recent_gerd = "1",
                                         recent_pepticulcer ="0",
                                         prior_gastric_cancer ="0",
                                         eth5 = "White",
                                         imd_person = "Least Deprived (1)"))

# Plot survival curves
plot(survfit7, col = "blue", lty = 1, xlim = c(0, 20), ylim = c(0, 1), xlab = "Time(years)", ylab = "Survival Probability", main = "50yo White Female,GERD,Least Deprived ")
lines(survfit8, col = "red", lty = 1)
legend("bottomleft", legend = c("PPI", "H2RA"), col = c("blue", "red"), lty = 1, cex = 0.7)


```

#### Sensitivity analysis

```{r}
# confounder selection With p value
# Recent peptic ulcer and GERD excluded
pvalmodel <- step(object = cox_multivariable, direction = "both", trace = 0, test = "Chisq")
summary(pvalmodel)

# confounder selection with AIC
AICmodel <- step(cox_multivariable)
summary(AICmodel)

#BMI included in final model
cox_sens1 <- coxph(Surv(survtime, died) ~ 
                            factor(ppi) + 
                            factor(recent_gerd) +
                            factor(recent_pepticulcer) + 
                            factor(prior_gastric_cancer) + 
                            factor(gender) + 
                            factor(eth5) + 
                            bmi+
                            factor(imd_person) +
                            age +
                            I(age^3),
                          data = analysis_dataset)
summary(cox_sens1)

# Stratification methods for selected confounders, gender and deprivation status
cox_sens2 <- coxph(Surv(survtime, died) ~ 
                          factor(ppi) + 
                          factor(recent_gerd) +
                          factor(recent_pepticulcer) + 
                          factor(prior_gastric_cancer) + 
                          strata(factor(gender)) + 
                          factor(eth5) + 
                          strata(factor(imd_person)) +
                          age +
                          I(age^3),
                        data = analysis_dataset)
summary(cox_sens2)



```
